\Large \bf Relationship of Weather and Maize Yields in Kenya

### 2018-11-14 writing/draft1/manuscript20.tex
adding

@@

Other studies have investigated the relationship of climate and yields in the context of climate change. \cite{MendelsohnDinar2000} have developed a simulation model to evaluate impacts of climate change on African agriculture. According to their study, agricultural GDP and climate change exhibit a hill-shaped relationship. Hence, global warming should be beneficial in high latitudes but harmful in low latitudes. The estimated loss of the Africa as continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}. \cite{kabubo2007} show that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial. Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}. \cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa. They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful. \cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania. The authors have shown that by 2050 the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather. \cite{Kabubo2015} have estimated effects of climate change on food security in Kenya. They found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped. Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
\cite{Mendelsohn2008} reviewed several studies on effects of climate change on agriculture in developing countries and based on the revision he confirmed that tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture. \cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia. The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions. According to another meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming. However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.

A number of important studies have focused on responses to drought related disasters or emergency situations. \cite{Sandstorm2017} have questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They have analysed causes of the food crisis identified by the humanitarian appeal documents. Majority of these documents have found food availability and food production to be the major causes of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts. According to \cite{Sandstorm2017}, a large share of humanitarian response budget has been focused on emergency food aid in comparison to the share focused on interventions to build resilience. The authors have suggested to increase the budget share focused on agricultural and livestock production to build resilience. Also other authors have promoted shifting from reactive to proactive approach in disaster risk management (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).  \cite{Nicholson2017} has recommended forecast based financing as it can avoid significant disaster losses. \cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses. Furthermore, risk assessment is essential for disaster risk management \citep{IPCC2012ch1}. In the light of these arguments, the main purpose of the present study is to develop a model for assessing the risk of drought based on weather and climate data. This should help to improve focus on proactive rather than reactive approach in drought disaster risk management.

@@

To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and agriculture on crop yields in Kenya. Hence, this is the first time that the effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country. To the best of our knowledge, this is also the first study \textcolor{blue}{which has utilised as comprehensive weather dataset as we have used: the frequency of our weather data was daily and the resolution:} \textcolor{red}{\textbf{!!! source of data}} 
% Kabubo-Mariara and Karanja (2015) similar study> but they use the more econometric approach (fixed effects models and random effects models as referred to by econometricians
\section{Methodology}\label{Method}

@@

Before beginning our research, we had to answer two essential % fundamental??
questions: `How should we measure food security?' and `How should we measure drought?' Taking into account previous literature and data availability, we considered three main approaches to measuring food security. These include using food prices, health and utility indicators or agricultural crop yields as measures of food security. Unfortunately, we had to exclude the health and utility indicators because of data availability issues. Although some health and utility indicators, such as Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (\citealp{CSI}) or the early warning phases are available in the NDMA Early Warning Bulletins, they are only available since mid $2013$. Furthermore, the scales do not seem to be consistent over all counties and frequently the values are not present at all. The relevant studies focused on food prices include \cite{Ubilava2018}, who has estimated relationship between sea surface temperature and commodity prices in tropics, \cite{robinson2010} who analysed implications of food production and price shocks welfare in Ethiopia or \cite{OxfamIDS} who have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM. \cite{brown2014book} and \cite{BROWN2015} have developed a quantitative models to investigate relationship of  weather and international price shocks on food prices in the developing world. \textcolor{red}{or the disadvantage of using pricess is..}We, however, decided not to use food prices as a measure of food security because prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless taking into account the spatial autocorrelation structure. Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and additional data which may not be available.  Therefore we opted for a \textcolor{red}{single} reduced form equation with crop yields as the dependent variable.  We have provided a broad overview of literature concerned with agricultural yields in Sections~\ref{Introduction} and further in this section.
We focused on maize as this crop is the principal staple food in Kenya and it is grown at 90\% of all farms in Kenya \citep{FAO}.

>>>>

Droughts and climate variability do not only affect agriculture but they have widespread consequences on food security in general. \cite{Ubilava2018} has found a connection between commodity prices and phases of the ENSO cycle. \cite{robinson2010} have used a Computable General Equilibrium (CGE) model for Ethiopia and they have shown that production shocks and resulting price increases have negative effects on farmers' income in drought-prone areas while moderate price increases can be beneficial for farmers in less drought-prone areas. \cite{OxfamIDS} have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM and they have estimated that climate change will result in decline in food productivity while food prices will increase. \cite{BROWN2015} have investigated food prices in $554$ local commodity markets in Africa, South Asia and Latin America and they concluded that almost 20\% of local market food prices had been affected by weather disturbances.

@@

Other studies have investigated the relationship of climate and yields in the context of climate change. \cite{MendelsohnDinar2000} have developed a simulation model to evaluate impacts of climate change on African agriculture. According to their study, agricultural GDP and climate change exhibit a hill-shaped relationship. Hence, global warming should be beneficial in high latitudes but harmful in low latitudes. The estimated loss of the Africa as continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}. \cite{kabubo2007} have shown that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial. Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}. \cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa. They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful. \cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania. The authors have shown that by 2050 the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather. \cite{Kabubo2015} have estimated effects of climate change on food security in Kenya. They found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped. Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
\cite{Mendelsohn2008} reviewed several studies on effects of climate change on agriculture in developing countries and based on the revision he confirmed that tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture. \cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia. The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions. According to another meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming. However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.

A number of important studies have focused on responses to drought related disasters or emergency situations. \cite{Sandstorm2017} have questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They have analysed causes of the food crisis identified by the humanitarian appeal documents. Majority of these documents have found food availability and food production to be the major causes of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts. According to \cite{Sandstorm2017}, a large share of humanitarian response budget has been focused on emergency food aid in comparison to the share focused on interventions to build resilience. The authors have suggested to increase the budget share focused on agricultural and livestock production to build resilience. Also other authors have promoted shifting from reactive to proactive approach in disaster risk management (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).  \cite{Nicholson2017} has recommended forecast based financing as it can avoid significant disaster losses. \cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses. Furthermore, risk assessment is essential for disaster risk management \citep{IPCC2012ch1}. In the light of these arguments, the main goal of the present study is to develop a model for assessing the risk of drought based on weather and climate data. This should help to improve focus on proactive rather than reactive approach in drought disaster risk management.

@@

To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya. Hence, this is the first time that effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country. To the best of our knowledge, this is also the first study \textcolor{blue}{which has utilised as comprehensive weather dataset as we have used: the frequency of our weather data was daily and the resolution:} \textcolor{red}{\textbf{!!! source of data}} 
% Kabubo-Mariara and Karanja (2015) similar study> but they use the more econometric approach (fixed effects models and random effects models as referred to by econometricians
\section{Methodology}\label{Method}

@@

Before beginning our research, we had to answer two essential % fundamental??
questions: `How should we measure food security?' and `How should we measure drought?' Taking into account the aim of this study, previous literature (see Section~\ref{Introduction}) and data availability, we were considering three main approaches to measuring food security. In particular, we considered using health and utility indicators, food prices or agricultural crop yields as measures of food security. Although some health and utility indicators including the Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (see \citealp{CSI}) or the NDMA early warning phases are available in the NDMA Early Warning Bulletins, the data are only available since July $2013$. Furthermore, the scales of indicators do not seem to be consistent over the counties and time and the values of indicators or even entire reports are frequently missing. Therefore, we decided not to use the health and utility indicators.

Using food prices as a measure of food security turned out be problematic as well. 
One problem is data availability as the food prices are usually available for market towns rather than at county level. Furthermore, the market data which are available in the publications of \cite{KNBS} (KNBS) only cover several years. Another problem is that prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless the model explicitly accounts for the spatial autocorrelation structure. Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and also additional data, which may not be available. Therefore we opted for agricultural crop yields as a measure of food security and we decided to estimate a \textcolor{red}{single} reduced form equation with crop yields as the dependent variable. In particular, we focus on maize yields as this crop is the principal staple food in Kenya and it is grown at 90\% of farms in Kenya \citep{FAO}.

### 2018-11-14 writing/draft1/manuscript3.tex
adding

\date{\normalsize{September 2018}}
\title{\Large \bf What Does the Weather Say about Yields in Kenya}
\author{Monika Novackova and Pedram Rowhani and Martin Todd}
\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}

@@

% table checked 3.10.2017

\singlespacing
\caption{\textit{\textbf{Climate change knowledge:} Multisplit lasso and jackknife OLS}}
% LassoKnow43se3.R      soubor Know435.R 
%musim lasso predelat. predtim jsem pouzivala lambda min. ale alex ma pravdu. je treba u vsech modelu pouzivat stejne pravidlo, takze lambda se.zmena 30.9.2017

%soubor Know435se.R
% finlit2 prestane byt signifikantni, takze vlastne se nic skoro nemeni, asi do prilohy tabulku s Know435se.R

\label{Know435} 
\centering
\begin{small}
\begin{tabular}{lclrcl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Gender = male&$<2\times 10^{-8}$&$^{***}$&$0.733$&$<2\times10^{-8}$&$^{***}$\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=0$~\tnote{a}&$0.038$&$^{*}$ &\multicolumn{3}{c}{\textit{Not included - base cat.}}\\
  \\
\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.000$& &$2.071$&$1.000$&\\
  \\
\vspace{-0.2cm}Cognitive reflection $=1$&$1.000$& &$0.283$&$0.129$& \\
  \\
\vspace{-0.2cm}Cognitive reflection $=1.5$&$1.000$& &$0.895$&$1.000$ \\
  \\
\vspace{-0.2cm}Cognitive reflection $=2$&$1.000$& &$0.628$&$1\times 10^{-5}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2.5$&$1.000$& &$1.098$&$1.000$&\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=3$&$0.046$&$^{*}$&$1.033$&$<2\times10^{-8}$&$^{***}$\\
\\
\vspace{-0.2cm}Financial literacy total score $=0.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\vspace{-0.2cm}Financial literacy total score $=1$&$1.000$&& \multicolumn{3}{c}{\textit{Not included}} \\
  \\
\vspace{-0.2cm}Financial literacy total score $=1.5$&$1.000$&&\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\vspace{-0.2cm}Financial literacy total score $=2$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\vspace{-0.2cm}Financial literacy total score $=2.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
  \\
  \vspace{-0.2cm}Financial literacy total score $=3$&$2\times10^{-5}$&$^{***}$&\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\hline
\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
    \item[a]The estimate is negative for cognitive reflection $=0$ while it is positive for cognitive reflection $=3$ in this model.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

>>>>

\date{\normalsize{October 2018}}
\title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens}
\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}

@@

\singlespacing
\caption{\textit{\textbf{Relationship of weather and maize yield:} Mixed models}}
% Peggy45,    can be found in foodSystems/Rcodes/Lags/..

\label{Peggy45} 
\centering
\begin{small}
\begin{tabular}{lclrlcl} 
\hline \vspace{-0.2cm} \\
  &\multicolumn{6}{c}{\textit{Estimated coefficients and p-values}}  \\

\vspace{-0.2cm} \\

  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects}}  &\multicolumn{2}{c}{{\textbf{Linear - unscaled}}}& \multicolumn{2}{c}{{\textbf{Linear-scaled}}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\

 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$-0.059$& $0.646$&\\
  \\
\vspace{-0.2cm}Prec. cum. MAM + OND lag 1, east&$0.054$& $0.041^{*}$&\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$& $0.477$&\\
  \\
  \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, east&$-0.045$& $0.244$&\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, west&$-0.045$& $0.244$&\\
  \\

      \vspace{-0.2cm}Prec. max OND, east&$0.0.133$& $0.013^{*}$&\\
  \\
        \vspace{-0.2cm}Prec. max OND, west&$0.144$& $0.004^{**}$&\\
  \\
    \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, east&$0.049$& $0.199$&\\
  \\
      \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$\\

  \\
\hline
\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$584$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} \item a comment
    \item[a]another comment
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

### 2018-11-14 writing/draft1/manuscript5.tex
adding

\subsection{Climate change knowledge}\label{ResKnow}

@@

\singlespacing
\caption{\textit{\textbf{Relationship of weather and maize yield:} Mixed models}}
% Peggy45,    can be found in foodSystems/Rcodes/Lags/..

\label{Peggy45} 
\centering
\begin{small}
\begin{tabular}{lclrlcl} 
\hline \vspace{-0.2cm} \\
  &\multicolumn{6}{c}{\textit{Estimated coefficients and p-values}}  \\

@@

  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects}}  &\multicolumn{2}{c}{{\textbf{Linear - unscaled}}}& \multicolumn{2}{c}{{\textbf{Linear-scaled}}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\

 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$-0.059$& $0.646$&\\
  \\
\vspace{-0.2cm}Prec. cum. MAM + OND lag 1, east&$0.054$& $0.041^{*}$&\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$& $0.477$&\\
  \\
  \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, east&$-0.045$& $0.244$&\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, west&$-0.045$& $0.244$&\\
  \\

      \vspace{-0.2cm}Prec. max OND, east&$0.0.133$& $0.013^{*}$&\\
  \\
        \vspace{-0.2cm}Prec. max OND, west&$0.144$& $0.004^{**}$&\\
  \\
    \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, east&$0.049$& $0.199$&\\
  \\
      \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$\\

  \\
\hline
\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$584$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} \item a comment
    \item[a]another comment
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

>>>>

@@

\singlespacing
\caption{\textit{\textbf{Linear mixed effects models:} Maize yield and weather}}
% Peggy45 & Peggy453   can be found in foodSystems/Rcodes/Lags/..

\label{Peggy45} 
\centering
\begin{small}
\begin{tabular}{lrccc} 
\hline \vspace{-0.2cm} \\

  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{2}{c}{\textbf{Scaled}}& \multicolumn{2}{c}{{\textbf{Unscaled\tnote{a}}}} \\

    \multicolumn{1}{l}{\vspace{0.1cm}}  &\textit{Estimate}&\textit{p-value}&\multicolumn{1}{c}{\textit{Estimate}}& \multicolumn{1}{c}{\textit{p-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$-0.059$&$0.646$&$1.379$&$1\times10^{-7}$$^{***}$\\
  \\
\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.054$&$0.041^{*}$&$3\times10^{-4}$&$0.177$\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$&$0.477$&$3\times10^{-4}$&$0.197$\\
  \\
  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.045$&$0.244$&$-0.029$&$0.001^{**}$\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.130$&$0.0001^{***}$&$-0.027$&$0.001^{**}$\\
  \\

      \vspace{-0.2cm}Prec. max OND, east&$0.133$&$0.013^{*}$&$0.004$&$0.085^{\bullet}$\\
  \\
        \vspace{-0.2cm}Prec. max OND, west&$0.144$&$0.004^{**}$&$0.011$&$9\times10^{-6}$$^{***}$\\
  \\
    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.049$&$0.199$&$0.075$&$0.369$\\
  \\
      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$&$0.494$&$7\times10^{-12}$ $^{***}$\\
  \\
  \hline
\vspace{-0.3cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.3cm}
\\
\hline
Intercept\\
Prec. cum. MAM+OND lag 1\\
 Prec. max OND\\
Temp. avg. Mar.-Sep. lag 1
  \\
  \hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$584$ observations; \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} 
    \item[a] The variety with unscaled variables fails to converge.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather}}
% Peggy45ln can be found in foodSystems/Rcodes/Lags/..

\label{Peggy45ln} 
\centering
\begin{small}
\begin{tabular}{lrlclcl} 
\hline \vspace{-0.2cm} \\

@@

  \multicolumn{1}{l}{\vspace{0.1cm}}  &\multicolumn{4}{c}{\textbf{Linear}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\

  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects}}  &\multicolumn{2}{c}{{\textbf{scaled}}}& \multicolumn{2}{c}{{\textbf{unscaled\tnote{a}}}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\
 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$-0.059$&$0.646$&$1.379$&$1\times10^{-7}$$^{***}$&$0.158$&$0.086^{\bullet}$\\
  \\
\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, e.&$0.054$&$0.041^{*}$&$3\times10^{-4}$&$0.177$&$0.066$&$0.008^{**}$\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM lag 1, w.&$0.025$&$0.477$&$3\times10^{-4}$&$0.197$&$-0.006$&$0.861$\\
  \\
  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, e.&$-0.045$&$0.244$&$-0.029$&$0.001^{**}$&$-0.036$&$0.292$\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, w.&$-0.130$&$0.0001^{***}$&$-0.027$&$0.001^{**}$&$-0.081$&$0.008^{**}$\\
  \\

      \vspace{-0.2cm}Prec. max OND, e.&$0.133$&$0.013^{*}$&$0.004$&$0.085^{\bullet}$&$0.081$&$0.056^{\bullet}$\\
  \\
        \vspace{-0.2cm}Prec. max OND, w.&$0.144$&$0.004^{**}$&$0.011$& $9\times10^{-6}$$^{***}$&$0.108$&$0.009^{**}$\\
  \\
    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, e.&$0.049$&$0.199$&$0.075$&$0.369$&$0.101$&$0.003^{**}$\\
  \\
      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, w.&$0.225$& $4\times10^{-12}$$^{***}$&$0.494$&$7\times10^{-12}$$^{***}$&$0.142$&$6\times10^{-7}$ $^{***}$\\
  \\
  \hline
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.2cm}
\\
\hline
\\
  \vspace{-0.2cm}Intercept\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM+OND lag 1\\
  \\
  \vspace{-0.2cm}Prec. max OND\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1\\

  \\
  \hline
\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$584$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} \item a comment
    \item[a] The variety with unscaled variables fails to converge.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}
